### 2014-03-28 LaTeX/RhodiumLit.tex
Added literature

varying hypotheses about the effect of war on economic development, war ruin v war renewal hypotheses

literature on sub-state determinants of conflict and outcome (pierskala hollenbach 2013)

	explain how what has been done in the spatial conflict literature and even the sub-nat/spat literature does not capture what we have done in this paper

% literature on endogeneity of conflict and economic performance (sambanis 2001, worldbank 2003)

% 	do we want to wade in to the endogeneity debate? i'm not sure we do that's why i've commented this out temporarily. seems to me our life would be made easier if we just tied ourselves to teh literature that puts econ growth as the dv nad conflict as the iv. u probably know better than me if this position is untenable.

@@

Not only do several studies link civil war to domestic economic performance, there is also evidence that civil wars have regional economic consequences.  \cite{murdoch:sandler:2002} find some evidence that states neighboring civil war states are more likely to experience poor short-term economic performance.  They attribute this effect to the disruption of trade and uncertainty about the potential for conflict to spread across the border.

@@

Much work has been done on the causal effects of economic performance on civil war.  Indeed, there is likely an endogenous relationship between economic performance and civil war; each exacerbates the other.  While our work here sidesteps this argument by focusing exclusively on observations of civil war, we will briefly review the relevant literature here.  In a report for the World Bank by \cite{collier:etal:2003}, the authors describe what they term the \textit{conflict trap}.  States that find themselves in the \textit{conflict trap} are those that have experienced civil war with, are subsequently affected by its economic and social consequences, and are therefore more likely to experience further civil conflict.  During civil wars, resources are diverted from productive economic activity to destructive activity.  These diverted resources act to stall progress during the conflict and are often used to destroy the infrastructure necessary for growth afterwards.  These changes to economic performance, as well as structural changes to the economy itself, make the resurgence of war more likely.  In accord with this theory, \cite{fearon:laitin:2003} aruge that poor economic growth is the primary condition conducive for civil war.  More specifically, they believe that strong economic growth proxies for robust governance and that states with low GDP growth likely have infrastructures that are unable to implement counterinsurgent policies.  In an effort to parse out the causal effect of economic growth shocks on civil war, \cite{miguel:etal:2004} instrument income growth with rainfall.  They find that rainfall is strongly correlated with income in sub-Saharan Africa, a region also prone to civil conflict in recent decades.   Using a two-stage estimation approach, they find that rainfall, their exogenous instrument for income, is positively correlated with the likelihood of civil war.

>>>>

@@

Not only do several studies link civil war to domestic economic performance, there is also evidence that civil wars have regional economic consequences.  \cite{murdoch:sandler:2002} find some evidence that states neighboring civil war states are more likely to experience poor short-term economic performance.  They attribute this effect to the disruption of trade and uncertainty about the potential for conflict to spread across the border.  In a follow up study, \cite{murdoch:sandler:2002} suggest that the spatial dispersion of economic effects from civil conflict differ from region to region.

@@

Much work has been done on the causal effects of economic performance on civil war.  Indeed, there is likely an endogenous relationship between economic performance and civil war; each exacerbates the other.  While our work here sidesteps this argument by focusing exclusively on observations of civil war, we will briefly review the relevant literature here.  In a report for the World Bank by \cite{collier:etal:2003}, the authors describe what they term the \textit{conflict trap}.  States that find themselves in the \textit{conflict trap} are those that have experienced civil war with, are subsequently affected by its economic and social consequences, and are therefore more likely to experience further civil conflict.  During civil wars, resources are diverted from productive economic activity to destructive activity.  These diverted resources act to stall progress during the conflict and are often used to destroy the infrastructure necessary for growth afterwards.  These changes to economic performance, as well as structural changes to the economy itself, make the resurgence of war more likely.

In accord with this theory, \cite{fearon:laitin:2003} aruge that poor economic growth is the primary condition conducive for civil war.  More specifically, they believe that strong economic growth proxies for robust governance and that states with low GDP growth likely have infrastructures that are unable to implement counterinsurgent policies.  In an effort to parse out the causal effect of economic growth shocks on civil war, \cite{miguel:etal:2004} instrument income growth with rainfall.  They find that rainfall is strongly correlated with income in sub-Saharan Africa, a region also prone to civil conflict in recent decades.   Using a two-stage estimation approach, they find that rainfall, their exogenous instrument for income, is positively correlated with the likelihood of civil war.

\subsubsection{Disaggregating Civil Wars}

Recently, scholars have begun to spatially disaggregate civil conflicts.  New data allows researchers to focus on how the geography of internal conflict varies.  \cite{pierskalla:hollenbach:2013} use subnational data on African states to asses the role cell phone coverage plays in facilitating violent conflict.  They theorize that cell phone coverage will enhance the collective action capabilities of rebel groups by improving coordination, communication, and in-group monitoring.  A series of empirical tests confirm this hypothesis and indicate that cell phone coverage corresponds to a 50\%-300\% increase in conflict likelihood for a given area (depending on the estimation strategy used).

\cite{berman:couttenier:2013} explore another sub-state determinant of civil conflict.  Recognizing that economic shocks are associated with changes to the probability of civil conflict,\footnote{For more, see \cite{miguel:etal:2004} and \cite{dube:vargas:forthcoming}.} they seek to determine where conflict will emerge when these shocks occur.  Given an external economic shock in a trading partner, Berman and Couttenier expect that states should be at an increased risk of experiencing conflict.  However, not all locations within a state will feel the effects equally.  Those areas most directly connected to the trading partner will be more likely to experience violent conflict than those areas that are less dependent on the trading partner.  They operationalize this measure of dependence, or ``remoteness,'' as distance from a seaport.  Indeed, they find that conflicts are more likley to arise near seaports following an economic shock than they are further away.
